---
title: "Manuscript Simulations"
author: "Claire Powers"
date: "2024-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r start, include=FALSE}

rm(list=ls())
library(tidyverse)
# library(EnvStats)
# library(RColorBrewer)
# library(ggthemes)
# library(iterators)
library(parallel)
library(doParallel)
library(foreach)

source("VR.Functions/ParamVals.R")
source("VR.Functions/subpop_mx_fun.R")
source("VR.Functions/logit2prob.R")
source("VR.Functions/survival_fun.R")
source("VR.Functions/growth_fun.R")
source("Sim.Functions/simfun_exp.R")
source("Sim.Functions/moveMx_fun.R")
seeds_df = read_csv("VR.Functions/fecundity.csv")
```

### Scenario 1: 
- symmetrical movement
- 2 subpops on cold & warm slopes
- Exponential growth
```{r ExpGrowth_sims_scen1}
## Simluation settings
reps=1
burn.in.yrs=100
sim.yrs=100
asp.effects = c(-1,1,1)
n.subpops=length(asp.effects)

## Variables to loop over
clim.sds = c(0.01,1,seq(5,20,by=5))
asp.mags = seq(0,20,by=5)

## Creating all combinations of movement rates
subpop.names = LETTERS[1:n.subpops] # Name subpops with letters
move.rates = seq(0,0.2,by=0.1)
move.dirs = expand.grid(from=subpop.names,to=subpop.names) %>% # get all combos
  # Remove rows that imply no movement (A to A)
  filter(from!=to) %>% 
   # Concatenate
  mutate(col.names = paste0(from,"to",to)) %>% 
  # save last column as a vector
  arrange(col.names) %>% pull() 

# Create movement data.frame to loop over
move.rates.df = matrix(data=rep(move.rates,length(move.dirs)),ncol=length(move.dirs)) %>% # generate matrix with movement rates
  as.data.frame() %>% 
  # Make all combos of movement rates
  expand.grid() %>% 
  filter(if_all(starts_with("V"),~ .==V1))
names(move.rates.df)=move.dirs # rename columns to make movement direction clear
rm(subpop.names,move.dirs,move.rates) # remove extra things from env
```

### Scenario 2: 
- **A**symmetrical movement
- 2 subpops on cold & warm slopes
- Exponential growth

#### S2 preparation 
```{r ExpGrowth_sims_scen2}
## Simluation settings
reps=1
burn.in.yrs=100
sim.yrs=100
asp.effects = c(-1,1)
n.subpops=length(asp.effects)

## Variables to loop over
clim.sds = c(0.01,1,seq(5,20,by=5))
asp.mags = seq(0,20,by=5)

## Creating all combinations of movement rates
subpop.names = LETTERS[1:n.subpops] # Name subpops with letters
move.rates = seq(0,0.2,by=0.1)
move.dirs = expand.grid(from=subpop.names,to=subpop.names) %>% # get all combos
  # Remove rows that imply no movement (A to A)
  filter(from!=to) %>% 
   # Concatenate
  mutate(col.names = paste0(from,"to",to)) %>% 
  # save last column as a vector
  arrange(col.names) %>% pull() 

# Create movement data.frame to loop over
move.rates.df = matrix(data=rep(move.rates,length(move.dirs)),ncol=length(move.dirs)) %>% # generate matrix with movement rates
  as.data.frame() %>% 
  # Make all combos of movement rates
  expand.grid() 
names(move.rates.df)=move.dirs # rename columns to make movement direction clear
rm(subpop.names,move.dirs,move.rates) # remove extra things from env
```

#### Run S2
```{r ExpGrowth_sims_c2}
cores = detectCores()
cl = makeCluster(cores/5)
registerDoParallel(cl = cl)

out_df <- foreach(i = 1:reps, .combine = rbind,.packages=c("tidyverse","EnvStats")) %:% 
  foreach(move.rates.row = 1:nrow(move.rates.df), .combine = rbind,.packages=c("tidyverse","EnvStats")) %:%
    foreach(clim.sd.i = clim.sds, .combine = rbind,.packages=c("tidyverse","EnvStats")) %:%
      foreach(asp.mag.i = asp.mags, .combine = rbind,.packages=c("tidyverse","EnvStats")) %dopar% {
        
        ##### Pull seed value from seed_df
        seeds.i = mean(seeds_df$seeds[seeds_df$asp.mag==asp.mag.i]) # This probbaly needs to move to inside the subpop function
        
        ##### This will go away - Just need it for testing out the loops
        move.rates.i = move.rates.df[move.rates.row,]
        
        ##### Create a movement matrix from the move.rates row
        move.mx.i = moveMx_fun(n.subpops,move.rates.i)
        
        df = simfun_exp(burn.in.yrs=burn.in.yrs, sim.yrs=sim.yrs, n.subpops=n.subpops,asp.effects = asp.effects,
                            clim.sd=clim.sd.i, asp.mag=asp.mag.i, seeds=seeds.i,move.mx = move.mx.i)
        tmp.df = df %>%
          filter(year>burn.in.yrs) %>% # Filter out burn-in years. This could also happen within the function.
          summarise(clim.mean = mean(clim),
                    clim.var = var(clim),
                    pop.var = var(pop.lam),
                    pop.lam = geoMean(pop.lam)) %>% 
          mutate(clim.sd = clim.sd.i,
                 # move.rate = move.rate.i,
                 asp.mag = asp.mag.i,
                 sim = i) %>% 
          merge(.,move.rates.i)
        return(tmp.df)
      
        # rm(df,tmp.df,clim.sd.i,asp.mag.i,move.rate.i)
        
      }
stopCluster(cl)
write_csv(out_df,paste0("server_output/output_",round(Sys.time()),".csv"))


```

asp.effect = 1
  pop.mx = matrix(data = 0,nrow=n.subpops*3, ncol=n.subpops*3)
  for(n in 1:n.subpops){
    
    # Add subpop asp.effect to parameters
    surv_params$asp.effect = growth_params$asp.effect = asp.effect
    
    # Create sub-population matrix
    subpop.mx = subpop_mx_fun(surv_params = surv_params,
                              growth_params = growth_params,
                              seeds = 1000)
    
    
    
    
    # put subpopulation MX into full population MX
    start.cell = start.cells[n]
    pop.mx[start.cell:(2+start.cell),start.cell:(2+start.cell)]=subpop.mx
  
    rm(subpop.mx)
  } # End subpop loop
  pop.mx = round(pop.mx,2)

  m.pop.mx = move.mx.m%*%pop.mx
  m.pop.mx[start.cells,start.cells]=pop.mx[start.cells,start.cells]
  
