---
title: "Plotting simulation results"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pacakges_data, include=FALSE}
library(tidyverse)
library(pals)
library(randomcoloR)
library(scico)

# rm(list=ls())

s1 <- read_csv("SimOutput/S1_2024-12-17 13:05:29.csv") %>% 
  #rbind(.,read_csv("SimOutput/S1_2024-12-17 18:31:40.csv")) %>% 
  rbind(.,read_csv("SimOutput/S1_2024-12-18 12:35:03.csv"))
s1_plot <- s1 %>% 
group_by(clim.sd,asp.mag,move.rate = AtoB,rho) %>% 
  summarize(agg.pop.var = mean(pop.var),
            mean.corr = mean(sub.corr,na.rm = T),
            pop.lam = mean(lt.pop.lam),
            .groups="drop") %>% 
  mutate(asp.effects = c("-1, +1")) %>% 
  group_by(asp.mag,clim.sd) %>% 
  mutate(move.inc = (pop.lam - pop.lam[move.rate==0])/pop.lam[move.rate==0],
         move.var.dec = (agg.pop.var[move.rate==0]-agg.pop.var)/agg.pop.var[move.rate==0])

s2 <- read_csv("SimOutput/S2_2024-12-17 23:33:35.csv") %>% 
  rbind(.,read_csv("SimOutput/S2_2024-12-18 15:06:16.csv"))
s2_plot <- s2 %>% 
group_by(clim.sd,asp.mag,move.rate = AtoB,rho) %>% 
  summarize(agg.pop.var = mean(pop.var),
            mean.corr = mean(sub.corr,na.rm = T),
            pop.lam = mean(lt.pop.lam)) %>% 
  mutate(asp.effects = c("-2, +1")) %>% 
  group_by(asp.mag,move.rate,clim.sd) %>% 
  mutate(move.inc = (pop.lam - min(pop.lam))/min(pop.lam)*100 )
```

```{r plot_theme, include=FALSE}

plot_theme <- function() {
  theme(
    panel.border = element_rect(colour = "black", fill = NA, linetype = 1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill="white",color="black"),
    panel.grid.major.y = element_line(colour = "gray80", linetype = 3, linewidth = 0.5),
    panel.grid.minor.y = element_blank(),# element_line(colour = "gray90", linetype = 3, linewidth = 0.5),
    panel.grid.major.x =  element_line(colour = "gray80", linetype = 3, linewidth = 0.5),
    panel.grid.minor.x = element_blank(),# element_line(colour = "gray90", linetype = 3, linewidth = 0.5)
    axis.text = element_text(size=12),
    axis.title = element_text(size=13),
    legend.text = element_text(size=12),
    legend.title = element_text(size=13),
    title = element_text(size=14)
  )
}

binwidth = 0.1
contour_colors <- function(df,binwidth){
  
  limits <- round(max(abs(1 - df$pop.lam)) * c(-1, 1) + 1,1)
  breaks <- seq(min(limits),max(limits),by=binwidth)
  pal <- scico(n = length(breaks),palette = "vik",direction = -1)
  
  ### Code for different options ###
  # colorRampPalette( brewer.pal(9,"RdYlGn") )(length(breaks))
  # brewer.spectral(n = length(breaks))
  
  }

```

---

### Scenario 1
    - Symmetrical movement between sub-populations
    - Symmetrical aspect effects
    - Exponential growth
 
#### Looking at basic dynamics
  - climate autocorrelation = 0.25
  - movement rate = 0.1
```{r S1_filter}
rho1 = 0
move.rates = c(0,0.05,0.1)
asp.mags = seq(0,20,4)

s1_plot2 <- s1_plot %>% 
  filter(move.rate%in%move.rates & rho == rho1&asp.mag%in%asp.mags)
```

---

#### *How do movement, climate standard deviation, and micro.habitat effects interact?*
 - This is assuming no autocorrelation in climate
```{r S1.mean.lam, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
ggplot(s1_plot2,aes(x=clim.sd,y=pop.lam,color=factor(move.rate*100)))+
  geom_hline(yintercept = 1,color="grey60")+
  geom_line(method = "loess",alpha=0.5)+
  geom_point()+
  facet_wrap(~asp.mag,nrow=2)+
  plot_theme()+
  labs(x = "Climate standard deviation", y = "Long-term population lambda",
       color="Movement\nrate (%)",title="Long-term lambda by climate standard deviation (x-axis),\nmovement rates (line colors), and microhabitat effects (sub-panels)")

```
**Movement between microclimates increases peak lambda values when microhabitat effects are strong.**

---

```{r S1.mean.lam2, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
ggplot(s1_plot2,aes(x=clim.sd,y=pop.lam,color=factor(asp.mag)))+
  geom_hline(yintercept = 1,color="grey60")+
  geom_line(stat="smooth",method="loess",alpha=0.5,linewidth=1.1)+
  geom_point()+
  facet_wrap(~move.rate,nrow=1)+
  plot_theme()+
  labs(x = "Climate standard deviation", y = "Long-term population lambda",
       color="Microhabitat effect",title="Long-term lambda by climate standard deviation (x-axis),\nmovement rates (subpanels), and microhabitat effects (line colors)")
```
**This is the same data shown in the plot above, but with the representation of movement rates and microhabitat effects swapped. Not sure it makes anything clearer**


#### How much does movement increase lambda for different aspect effects, relative to scenarios with no movement?

```{r S1.mean, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}

ggplot(filter(s1_plot2,move.rate!=0),aes(x=clim.sd,y=move.inc*100,color=factor(asp.mag))) +
  geom_point()+
  geom_line(stat = "smooth",method="loess",formula = y~x,linewidth=1.1,alpha=0.5)+
  facet_wrap(~(move.rate*100))+
  labs(x="Climate standard deviation",
  y="% increase in lambda
       relative to scenarios with no movement",
  color="Microhabitat effect")+
  plot_theme()
```

**This is showing the % increase in l.t. lambda values for given climate S.D.-micro-habitat effect--movement rate (5% and 10%) combinations. An interesting factor here is that the highest % increases are different for different microhabitat effects (makes sense) and that % increases are positively correlated with the strength of microhabitat effeccts**

---

```{r S1.var, fig.height=5, fig.width=10}
ggplot(filter(s1_plot2,clim.sd%in%c(0:20)),aes(x=clim.sd,y=agg.pop.var,color=factor(move.rate*100)))+
  geom_point()+
  geom_line(alpha=0.5)+
  facet_wrap(~asp.mag,nrow=2)+
  plot_theme()+
  labs(x = "Climate standard deviation", y = "Variance in long-term population lambda",
       color="Movement\nrate (%)",
       title = "VARIANCE in lambda by climate standard deviation (x-axis), \nmovement rates (line colors), and microhabitat effects (sub-panels)")
```

\
 **Variance in long-term lambda increases with the strength of micro-habitat effects. For a given microhabitat effect, movement between micro-climates reduces variance. It's unclear whether the variance reduction scales with **

\

--- 

#### *How does climate autocorrelation change this?*
```{r S2.clim.autocorr.lambda}
move.rates = c(0,0.05)
asp.mags = seq(0,20,4)
rho1 <- c(0,0.25,0.5)

s1_plot2 <- s1_plot %>% filter(move.rate%in%move.rates & asp.mag%in%asp.mags & rho %in% rho1)

ggplot(s1_plot2,aes(x=clim.sd,y=pop.lam,color=factor(rho)))+
  geom_hline(yintercept = 1,color="grey30")+
  geom_line(stat="smooth",method="loess",alpha=0.75, formula = y~x)+
  geom_point(size=0.7)+
  facet_grid(move.rate~asp.mag)+
  plot_theme()+
  labs(x="Climate SD",y="Long-term lambda",color="Clim. \nAutocorr.",title="Lambda by climate SD, 
       microhabitat effects (columns),
       and movement rates (rows)")

```
 
 - If there is movement between microclimates (lower row), autocorrelation in climate increases long-term lambda values when microhabitat effects are high. Without movement it looks like autocorrelation has essentially no effect on long-term lambda for any microhabitat effect strengths.
\
#### How does climate autocorrelation change variance in lambda?

```{r S1.clim.autocorr.var, fig.width=10}
move.rates = c(0,0.05)
asp.mags = seq(0,20,4)

s1_plot2 <- s1_plot %>% filter(move.rate%in%move.rates & asp.mag%in%asp.mags & rho %in% rho1)

ggplot(s1_plot2,aes(x=clim.sd,y=agg.pop.var,color=factor(rho)))+
  geom_line(stat="smooth",formula=y~x,method="loess")+
  geom_point()+
  facet_grid(move.rate~asp.mag)+
  plot_theme()+
  labs(title="Variance in lambda values by:
       - climate standard deviation (x axis),
       - climate auto-correlation (line colors),
       - microhabitat effects (sub-plot columns),
       - and movement rates (sub-plot rows)",
       x="Climate SD", y="Variance in lambda",color="Clim. \nAutocorr.")
```
\
  **More autocorrelation in climate values increases variance in lambda. Movement still reduces variance.**


#### Now looking at correlation between subpopulation annuaal lambdas
```{r S1.sp.corr1}
move.rates <- c(0.025)
asp.mags <- seq(0,20,2)
rho1 <- c(0.25)

s1_plot2 <- s1_plot %>% filter(move.rate%in%move.rates & asp.mag%in%asp.mags & rho%in%rho1)

ggplot(s1_plot2,aes(x=clim.sd,y=mean.corr,color=factor(asp.mag)))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d()+
  plot_theme()
```
\
#### What is causing the hump shape in the plot above?
```{r subcorr.deepdive, eval=FALSE, include=FALSE}

S1_out_df <- S1_out_df %>% mutate(mean.ann.lam = (s1+s2)/2)

ggplot(filter(S1_out_df,year%in%200:210 & clim.sd%in%4:12 & asp.mag %in% c(10,15)),aes(x=year))+
  geom_hline(yintercept = 1)+
  geom_point(aes(y=s1,color="S1"))+
  geom_line(aes(y=s1,color="S1"))+ 
  geom_point(aes(y=s2,color="S2"))+
  geom_line(aes(y=s2,color="S2"))+
  facet_grid(asp.mag~clim.sd)+
  plot_theme()+
  theme(axis.text.x = element_blank())

ggplot(filter(S1_out_df,asp.mag==15),aes(x=clim))+
  geom_hline(yintercept = 1,color="red")+
  geom_point(aes(y=s1),color="blue")+
  geom_point(aes(y=s2),color="gold2")+
  facet_wrap(~clim.sd,scales="free")+
  plot_theme()

ggplot(filter(S1_out_df,asp.mag==15),aes(x=s1,y=s2))+
  geom_point()+
  geom_hline(yintercept = 1)+
  geom_vline(xintercept = 1)+
  facet_wrap(~clim.sd,scales="free")+
  plot_theme()

```

### Contour plots
```{r}

cols = contour_colors(s1_plot,binwidth = 0.05)

ggplot(s1_plot,aes(x=clim.sd,y=asp.mag,z=pop.lam))+
  geom_contour_filled()+
  facet_grid(rho~move.rate)+
  scale_fill_manual(values = cols,name="Lambda")+
  plot_theme()+
  labs(x="Climate SD", y="Microhabitat effect",
       title = "Columns = movement rates \nRows = Climate autocorrelation values")

```

```{r}
ggplot(s1_plot,aes(x=clim.sd,y=asp.mag,z=agg.pop.var))+
  geom_contour_filled()+
  facet_grid(rho~move.rate)+
  plot_theme()+
  labs(x="Climate SD", y="Microhabitat effect",
       title = "Columns = movement rates \nRows = Climate autocorrelation values")
```

```{r}
ggplot(filter(s2_plot),aes(x=clim.sd,y=asp.mag,z=pop.lam))+
  geom_contour_filled()+
  facet_grid(rho~move.rate)+
  scale_fill_manual(values = cols,name="Lambda")+
  plot_theme()+
  labs(x="Climate SD", y="Microhabitat effect",
       title = "Columns = movement rates \nRows = Climate autocorrelation values")


ggplot(filter(s1_plot),aes(x=clim.sd,y=asp.mag,z=pop.lam))+
  geom_contour_filled()+
  facet_grid(rho~move.rate)+
  scale_fill_manual(values = cols,name="Lambda")+
  plot_theme()+
  labs(x="Climate SD", y="Microhabitat effect",
       title = "Columns = movement rates \nRows = Climate autocorrelation values")


cowplot::plot_grid(s1_contour,s2_contour,nrow = 2)

```

