---
title: "Plotting simulation results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pacakges_data, include=FALSE}
library(tidyverse)
library(pals)
library(randomcoloR)

rm(list=ls())


s1 <- read_csv("SimOutput/S1_2024-12-12 13:37:18.csv")
s1_plot <- s1 %>% 
group_by(clim.sd,asp.mag,move.rate = AtoB,rho) %>% 
  summarize(agg.pop.var = mean(pop.var),
            mean.corr = mean(sub.corr,na.rm = T),
            pop.lam = mean(lt.pop.lam)) %>% 
  mutate(asp.effects = c("-1, +1"))

s2 <- read_csv("SimOutput/S2_2024-12-12 14:32:17.csv")
s2_plot <- s2 %>% 
group_by(clim.sd,asp.mag,move.rate = AtoB,rho) %>% 
  summarize(agg.pop.var = mean(pop.var),
            mean.corr = mean(sub.corr,na.rm = T),
            pop.lam = mean(lt.pop.lam)) %>% 
  mutate(asp.effects = c("-1, +1"))
```

```{r plot_theme, include=FALSE}

plot_theme <- function() {
  theme(
    panel.border = element_rect(colour = "black", fill = NA, linetype = 1),
    panel.background = element_rect(fill = "white"),
    strip.background = element_rect(fill="white",color="black"),
    panel.grid.major.y = element_line(colour = "gray90", linetype = 3, linewidth = 0.5),
    panel.grid.minor.y = element_line(colour = "gray90", linetype = 3, linewidth = 0.5),
    panel.grid.major.x =  element_blank(),
    panel.grid.minor.x = element_blank()
  )
}

binwidth = 0.1
contour_colors <- function(df,binwidth){
  
  limits <- round(max(abs(1 - df$pop.lam)) * c(-1, 1) + 1,1)
  breaks <- seq(min(limits),max(limits),by=binwidth)
  pal <- brewer.spectral(n = length(breaks))
  
  }

```

---

### Scenario 1
 - Symmetrical movement between sub-populations
 - Symmetrical aspect effects
 - Exponential growth
 
#### Basic dynamics
 - climate autocorrelation = 0.25
 - movement rate = 0.1
```{r S1_filter}
rho1 = 0
move.rates = c(0,0.04,0.1)
asp.mags = c(0,5,9,10,11,12,13,15)

s1_plot2 <- s1_plot %>% filter(move.rate%in%move.rates & rho == rho1&asp.mag%in%asp.mags)
```

---

#### *What are basic dynamics between movement, climate standard deviation, and micro.habitat effects?*
 - This is assuming no autocorrelation in climate
```{r S1.mean, fig.height=5, fig.width=10}
 ggplot(s1_plot2,aes(x=clim.sd,y=pop.lam,color=factor(move.rate*100)))+
  geom_line()+
  facet_wrap(~asp.mag,nrow=1)+
  plot_theme()+
  labs(x = "Climate standard deviation", y = "Long-term population lambda",
       color="Movement rate (%)",title="Long-term lambda by climate standard deviation (x-axis),\nmovement rates (line colors), and microhabitat effects (sub-panels)")
```

**- Movement between microclimates increases peak lambda values when microhabitat effects are strong.**

---

```{r S1.var, fig.height=5, fig.width=10}
ggplot(s1_plot2,aes(x=clim.sd,y=agg.pop.var,color=factor(move.rate*100)))+
  geom_line()+
  facet_wrap(~asp.mag,nrow=1)+
  plot_theme()+
  labs(x = "Climate standard deviation", y = "Variance in long-term population lambda",
       color="Movement rate (%)",
       title = "VARIANCE in lambda by climate standard deviation (x-axis), \nmovement rates (line colors), and microhabitat effects (sub-panels)")
```

\
 **- Variance in long-term lambda increases with the strength of micro-habitat effects. For a given microhabitat effect, movement between micro-climates reduces variance, and that reduction appears to scale with microhabitat effect strengths to some degree.**
\

--- 

#### *How does climate autocorrelation change this?*
```{r S2.clim.autocorr.lambda}
move.rates = c(0,0.04)
asp.mags = c(0,5,10,15)

s1_plot2 <- s1_plot %>% filter(move.rate%in%move.rates & asp.mag%in%asp.mags)

ggplot(s1_plot2,aes(x=clim.sd,y=pop.lam,color=factor(rho)))+
  geom_line()+
  facet_grid(move.rate~asp.mag)+
  plot_theme()
```
 
 - If there is movement between microclimates (lower row), autocorrelation in climate increases long-term lambda values when microhabitat effects are high. Without movement it looks like autocorrelation has essentially no effect on long-term lambda for any microhabitat effect strengths.
\
#### How does climate autocorrelation change this?

```{r S1.clim.autocorr.var, fig.width=10}
move.rates = c(0,0.04)
asp.mags = c(0,5,10,15)

s1_plot2 <- s1_plot %>% filter(move.rate%in%move.rates & asp.mag%in%asp.mags)

ggplot(s1_plot2,aes(x=clim.sd,y=agg.pop.var,color=factor(rho)))+
  #geom_point()+
  geom_line()+
  # geom_smooth()+
  facet_grid(move.rate~asp.mag)+
  plot_theme()+
  labs(title="Variance in lambda values by:
       - climate standard deviation (x axis),
       - climate auto-correlation (line colors),
       - microhabitat effects (sub-plot columns),
       - and movement rates (sub-plot rows)")
```
\
    - Interestingly, variance in lambdas is also higher with climate autocorrelation. 

```{r S1}
move.rates <- c(0.04)
asp.mags <- seq(10,20,1)
rho1 <- c(0.25)

s1_plot2 <- s1_plot %>% filter(move.rate%in%move.rates & asp.mag%in%asp.mags & rho%in%rho1)

ggplot(s1_plot2,aes(x=clim.sd,y=mean.corr,color=factor(asp.mag)))+
  geom_point()+
  geom_line()+
  scale_color_viridis_d()
```
\
    - The correlation between annual subpopulation lambdas decreases from 1 as microhabitat effects increase. The rate of decline is slower 

```{r}

ggplot(S1_out_df,aes(x=clim))+
  geom_histogram()+
  facet_wrap(~clim.sd)

```

