---
title: "Simulate population data"
author: "Claire Powers"
date: "2025-02-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
rm(list=ls())
library(tidyverse)
library(EnvStats)
library(parallel)
library(doParallel)
library(foreach)


seeds_df <- read_csv("seedSearchOutput/seedSearch_2025-01-29 21:30:11.114866.csv")

source("ModelFunctions/ParamVals.R")
source("SimFunctions/simData_fun.R")
source("SimFunctions/clim_vals_fun.R")
source("ModelFunctions/survival_fun.R")
source("ModelFunctions/growth_fun.R")
source("ModelFunctions/logit2prob.R")
source("ModelFunctions/subpop_mx_fun.R")


# Constants
timesteps <- 50
start_pop <- c(840,20,50)

# Things to eventually loop over
clim.sd <- 0.001
rho <- 0
asp.mag <- 0
asp.effects <- c(-1,1)
move.rate <- 0.05

# Within a loop later
clim_vals <- clim_vals_fun(clim.sd = clim.sd, rho=rho, timesteps = timesteps)
seed_vec <- seeds_df$seeds[seeds_df$asp.mag==asp.mag & seeds_df$asp.effects==paste(asp.effects,collapse = ",")]
seeds <- mean(seed_vec)

df <- simulate_pop_data(surv_params = surv_params,
                        growth_params = growth_params,
                        seeds = seeds,
                        start_pop = start_pop,
                        timesteps = 20,
                        clim_vals = clim_vals,
                        asp.effects = c(-1,1),
                        asp.mag = asp.mag)

```


### making a for-loop out of the code above
```{r}

reps <- 2
clim.sds <- seq(0,15,5);clim.sds[clim.sds==0]=0.01
asp.mags <- seq(0,15,5)
rhos <- seq(0,0.5,0.2)
move.rates <- seq(0,0.2,0.1)
asp.effects <- list(c(-1,1))
start_pop <- c(840,20,50)
timesteps <- 20

cl = makeCluster(spec = 2)
registerDoParallel(cl = cl)

sim_data_out <- foreach(rep.i = 1:reps, .combine = rbind,.packages=c("tidyverse","EnvStats")) %:% 
  foreach(asp.eff.i = 1:length(asp.effects), .combine = rbind,.packages=c("tidyverse","EnvStats")) %:%
    foreach(move.rate.i = move.rates, .combine = rbind,.packages=c("tidyverse","EnvStats")) %:%
      foreach(clim.sd.i = clim.sds, .combine = rbind,.packages=c("tidyverse","EnvStats")) %:%
        foreach(asp.mag.i = asp.mags, .combine = rbind,.packages=c("tidyverse","EnvStats")) %:% 
          foreach(rho.i = rhos,.combine = rbind,.packages=c("tidyverse","EnvStats")) %dopar% {
            
            asp.effects.i <- asp.effects[asp.eff.i] %>% unlist()
            
            clim_vals <- clim_vals_fun(clim.sd = clim.sd.i, rho=rho.i, timesteps = timesteps)
            seeds.i = mean(seeds_df$seeds[seeds_df$asp.mag==asp.mag.i & seeds_df$asp.effects == paste(asp.effects.i,collapse = ",")])
            
            df <- simulate_pop_data(surv_params = surv_params,
                        growth_params = growth_params,
                        seeds = seeds.i,
                        start_pop = start_pop,
                        timesteps = timesteps,
                        clim_vals = clim_vals,
                        asp.effects = asp.effects.i,
                        asp.mag = asp.mag.i) %>% 
              mutate(rep = rep.i,
                     asp.mag = asp.mag.i,
                     rho = rho.i,
                     clim.sd = clim.sd.i,
                     asp.effects.grp =asp.effects[asp.eff.i] )
            
            write_csv(df, paste0("SimOutput/simData/V2_",round(Sys.time()),".csv"))
            # write_csv(df,paste0("SimOutput/simData/clim",clim.sd.i,"_aspmag",asp.mag.i,"_aspeffs",paste(asp.effects.i,collapse = "."),"_rho",rho.i,"_r",rep.i,".csv"))
            return(df)
            
          }
stopCluster(cl)
```

```{r}

sim_data_fns <- list.files("SimOutput/simData/",full.names = T)
sim_data <- vroom::vroom(sim_data_fns) %>% 
  filter(timestep>0) %>% 
  filter(is.na(asp.mag))
```


### Looking at output to check it
```{r}
tallies <- df %>% 
  group_by(individual_id) %>% tally()

stage.nums <- df %>% 
  group_by(timestep,stage) %>% 
  summarise(size = sum(survival_status))

ggplot(stage.nums,aes(x=timestep,y=size,color=stage))+
  geom_point()

table(df$stage,df$timestep)

pop.size <- df %>% 
  group_by(timestep) %>% 
  summarise(size=sum(survival_status)) %>% 
  mutate(ann.lam = size/lag(size)) %>% 
  mutate(geo.mean = geoMean(ann.lam,na.rm=T))

ggplot(pop.size,aes(x=timestep,y=size))+
  geom_point()

ggplot(pop.size,aes(x=timestep,y=ann.lam))+
  geom_point()+
  geom_hline(yintercept = pop.size$geo.mean)

```

- Take the actual values that you have and say what are the mean and the variance of lambda for that model with two populations
- Key question: If you ignore subpopulations and you estimate the mean, variance, and covariance 
- Run the stochastic model: whats the dependency of each vital rate on climate. What's the unexplained variance for surv, g, repro,
- What is it's prediction about stochastic growth rate or extinction rate, etc. 

1. In each year, what is the estimated growth and surv or juvs
2. What's the mean and temporal vital rate and covariation
3. Function of year and aspect effect 
4. Assume you know the movement rate

50 adults and start at stable stage structure for the average climate
20 years 

## Getting SSD
```{r}
seeds
mean.mx <- subpop_mx_fun(surv_params = surv_params,growth_params = growth_params,seeds.in = 21)
mean.mx

Re(eigen(mean.mx)$values[1])

ssd <- eigen(mean.mx)$vectors[,1]/sum(eigen(mean.mx)$vectors[,1])
ssd.multiplier <- 50/ssd[3]

ssd.nums <- round(ssd*ssd.multiplier)

size_mx <- matrix(ncol=3,nrow=timesteps+1)
size_mx[1,] <- ssd.nums

for(i in 1:timesteps){
  
  growth_params$clim <- surv_params$clim <- rnorm(1)
  mean.mx <- subpop_mx_fun(surv_params = surv_params,growth_params = growth_params,seeds.in = 21)
  size_mx[i+1,] <- mean.mx%*%as.numeric(size_mx[i,])
  
}

size_df <- as.data.frame(size_mx) %>% 
  mutate(timesteps = 1:(timesteps+1),
         total = V1+V2+V3)

ggplot(size_df,aes(x=timesteps,y=total))+
  geom_point()
```






