---
title: "Simulate population data"
author: "Claire Powers"
date: "2025-02-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
rm(list=ls())
library(tidyverse)
library(data.table)
library(EnvStats)
library(parallel)
library(doParallel)
library(foreach)

seeds_df <- read_csv("seedSearchOutput/seedSearch_2025-02-27 20:15:48.624276.csv") %>% 
  mutate(seeds2 = floor(seeds)) %>% 
  group_by(asp.effects,asp.mag,clim.sd) %>% 
  summarize(avg.seeds = round(mean(seeds)))

source("ModelFunctions/ParamVals.R")
source("SimFunctions/simData_fun.R")
# source("SimFunctions/simData_fun2.R")
# source("SimFunctions/CopyOfsimData_fun.R")
source("SimFunctions/clim_vals_fun.R")
source("ModelFunctions/survival_fun.R")
source("ModelFunctions/growth_fun.R")
source("ModelFunctions/logit2prob.R")
source("ModelFunctions/subpop_mx_fun.R")
```

### Simulate data
Very basic:
- 2 subpopulations, aspect effects = -1,1
- optimal climate standard deviation
- Range of movement rates and 

```{r}

# Define your parameters
reps <- 5
clim.sds <- c(0.01,seq(1, 20, 1))
asp.mags <- c(0, 1, 5, 8:15)
rhos <- 0
move.rates <- c(0.05, 0.2, 0.4)
asp.effects <- c(-1, 1)
timesteps <- 20

# Create output directories
# sapply(asp.mags, function(x)
#   if (!dir.exists(paste0("SimOutput/simulatedData/V2/aspmag", x)))
#     dir.create(paste0("SimOutput/simulatedData/V2/aspmag", x))
# )

# Parameter combinations
param_grid <- expand.grid(
  rep = 1:reps,
  asp.mag = asp.mags,
  clim.sd = clim.sds,
  move.rate = move.rates,
  rho = rhos
)

# Set up parallel cluster
n.cores <- 5
cl <- makeCluster(n.cores)
registerDoParallel(cl)

# Limit nested threading
Sys.setenv(OMP_NUM_THREADS = "1")
Sys.setenv(MKL_NUM_THREADS = "1")
Sys.setenv(OPENBLAS_NUM_THREADS = "1")
Sys.setenv(NUMEXPR_NUM_THREADS = "1")

# Parallel loop over the parameter grid
invisible(
  foreach(i = 1:nrow(param_grid), .packages = c("tidyverse", "EnvStats", "data.table")) %dopar% {
    
    row <- param_grid[i, ]
    
    # Create output directories
    if(!dir.exists(paste0("SimOutput/simulatedData/V2/aspmag", row$asp.mag))){
      dir.create(paste0("SimOutput/simulatedData/V2/aspmag", row$asp.mag))
    }
    
    growth_params.i <- growth_params
    surv_params.i <- surv_params

    clim_vals <- clim_vals_fun(
      clim.sd = row$clim.sd,
      rho = row$rho,
      timesteps = timesteps
    )
    
    seeds.i <- seeds_df$avg.seeds[seeds_df$asp.mag == row$asp.mag & seeds_df$asp.effects == paste(asp.effects, collapse = ",")]
    
    df.i <- simulate_pop_data(
      surv_params = surv_params.i,
      growth_params = growth_params.i,
      seeds = seeds.i,
      timesteps = timesteps,
      clim_vals = clim_vals,
      asp.effects = asp.effects,
      move.rate = row$move.rate,
      asp.mag = row$asp.mag
    ) %>%
      mutate(
        rep = row$rep,
        rho = row$rho,
        clim.sd = row$clim.sd,
        move.rate = row$move.rate,
        asp.effects.grp = factor(paste(asp.effects, collapse = ", "))
      )
    
    fwrite(df.i, paste0("SimOutput/simulatedData/V2/aspmag", row$asp.mag,
                "/clim", row$clim.sd,
                "_rho", row$rho,
                "_move", row$move.rate,
                "_rep", row$rep,
                ".csv.gz"))
    
    rm(df.i, clim_vals, growth_params.i, surv_params.i, seeds.i)
    return(NULL)
  } 
)
# Stop the cluster
stopCluster(cl)
```

```{r simple_data_sim}

reps <- 5
clim.sds <- c(0.01,seq(1,20,1))
asp.mags <- c(0,1,8:15)
rhos <- 0 # c(0,0.2,0.2)
move.rates <- c(0.05,0.2,0.4,0.8)
asp.effects <- c(-1,1)# list(c(-1,1))
timesteps <- 20

sapply(asp.mags,function(x)
  if(!dir.exists( paste0("SimOutput/simulatedData/V2/aspmag",x) ))
    dir.create(paste0("SimOutput/simulatedData/V2/aspmag",x)) 
  )


cl = makeCluster(spec = reps*2)
registerDoParallel(cl = cl)

foreach(rep.i = 1:reps,.packages=c("tidyverse","EnvStats","data.table")) %:%
   foreach(asp.mag.i = asp.mags) %:% 
     foreach(clim.sd.i = clim.sds) %:%
       foreach(move.rate.i = move.rates) %:%
          foreach(rho.i = rhos) %dopar% {

            growth_params.i <- growth_params;surv_params.i <- surv_params
            asp.effects.i <-  asp.effects # asp.effects[asp.eff.i] %>% unlist()

            clim_vals <- clim_vals_fun(clim.sd = clim.sd.i, rho=rho.i, timesteps = timesteps)
            
            # plot(clim_vals);lines(clim_vals)
            seeds.i = mean(seeds_df$seeds[seeds_df$asp.mag==asp.mag.i & seeds_df$asp.effects == paste(asp.effects.i,collapse = ",")])

            df.i <- simulate_pop_data(surv_params = surv_params.i,
                        growth_params = growth_params.i,
                        seeds = seeds.i,
                        timesteps = timesteps,
                        clim_vals = clim_vals,
                        asp.effects = asp.effects.i,
                        move.rate = move.rate.i,
                        asp.mag = asp.mag.i) %>% 
              mutate(rep = rep.i,
                     asp.mag = asp.mag.i,
                     rho = rho.i,
                     clim.sd = clim.sd.i,
                     move.rate = move.rate.i,
                     asp.effects.grp = factor(paste(asp.effects.i,collapse=", ")) )
            
            # write_csv(df, paste0("SimOutput/simData/server/V1/",round(Sys.time()),".csv"))
            write_csv(df.i,paste0("SimOutput/simulatedData/V2/aspmag",asp.mag.i,"/clim",clim.sd.i,"_rho",rho.i,"_move",move.rate.i,"_rep",rep.i,".csv"))
            rm(df.i,clim_vals)
            return(NULL)
            
            }
  
stopCluster(cl)
```

### Looking at output to check it
```{r}

sim_data_fns <- list.files("SimOutput/simulatedData/V1/aspmag10",
                           pattern = "clim10_rho0_move0.05",
                           full.names = T)

sim_data <- vroom::vroom(sim_data_fns,
                         col_types = ) %>% 
  filter(timestep>0) %>% 
  filter(!is.na(asp.mag))

stage.nums <- sim_data %>% 
  filter(end.stage!="dead") %>% 
  group_by(timestep,stage = end.stage,rep,asp.effect) %>% 
  summarise(size = sum(survival_status,na.rm=T)) 

ggplot(stage.nums,aes(x=timestep,y=size,color=factor(rep)))+
  geom_point(aes(shape=factor(asp.effect)))+
  geom_line(aes(linetype=factor(asp.effect)))+
  facet_wrap(~stage,scales="free_y")+
  theme_few(base_size = 14)+
  labs(title = "rnorm")+
  theme(panel.grid.major = element_line(colour = "grey90",linetype=3),
        panel.grid.minor = element_line(colour = "grey90",linetype=3))

pop.size <- sim_data %>% 
  group_by(timestep,rep) %>% 
  summarise(size=sum(survival_status,na.rm=T)) %>% 
  mutate(ann.lam = size/lag(size)) %>% 
  mutate(geo.mean = geoMean(ann.lam,na.rm=T))

ggplot(pop.size,aes(x=timestep,y=size,color=factor(rep)))+
  geom_point()+
  geom_line()s

```

# Test chunk
```{r}
# Constants
timesteps <- 20
start_pop <- c(840,20,50)

# Things to eventually loop over
clim.sd <- 1
rho <- 0
asp.mag <- 2
asp.effects <- c(-1,1)
move.rate <- 0.1

# Within a loop later
clim_vals <- clim_vals_fun(clim.sd = clim.sd, rho=rho, timesteps = timesteps)
seed_vec <- seeds_df$seeds[seeds_df$asp.mag==asp.mag & seeds_df$asp.effects==paste(asp.effects,collapse = ",")]
seeds <- mean(seed_vec)

df1 <- simulate_pop_data(surv_params = surv_params,
                        growth_params = growth_params,
                        seeds = seeds,
                        start_pop = start_pop,
                        timesteps = timesteps,
                        clim_vals = clim_vals,
                        asp.effects = c(-1,1),
                        asp.mag = asp.mag,
                        move.rate = move.rate)


```


- Take the actual values that you have and say what are the mean and the variance of lambda for that model with two populations
- Key question: If you ignore subpopulations and you estimate the mean, variance, and covariance 
- Run the stochastic model: whats the dependency of each vital rate on climate. What's the unexplained variance for surv, g, repro,
- What is it's prediction about stochastic growth rate or extinction rate, etc. 

1. In each year, what is the estimated growth and surv or juvs
2. What's the mean and temporal variance and covariation
3. Function of year and aspect effect 
4. Assume you know the movement rate

50 adults and start at stable stage structure for the average climate
20 years 

## Getting SSD
```{r}
seeds
mean.mx <- subpop_mx_fun(surv_params = surv_params,growth_params = growth_params,seeds.in = 21)
mean.mx

Re(eigen(mean.mx)$values[1])

ssd <- eigen(mean.mx)$vectors[,1]/sum(eigen(mean.mx)$vectors[,1])
ssd.multiplier <- 50/ssd[3]

ssd.nums <- round(ssd*ssd.multiplier)

size_mx <- matrix(ncol=3,nrow=timesteps+1)
size_mx[1,] <- ssd.nums

for(i in 1:timesteps){
  
  growth_params$clim <- surv_params$clim <- rnorm(1)
  mean.mx <- subpop_mx_fun(surv_params = surv_params,growth_params = growth_params,seeds.in = 21)
  size_mx[i+1,] <- mean.mx%*%as.numeric(size_mx[i,])
  
}

size_df <- as.data.frame(size_mx) %>% 
  mutate(timesteps = 1:(timesteps+1),
         total = V1+V2+V3)

ggplot(size_df,aes(x=timesteps,y=total))+
  geom_point()
```






