---
title: "Setting parameter values"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r start, message=TRUE, warning=TRUE, include=FALSE}
library(tidyverse)
library(EnvStats)
library(optimx)

rm(list=ls())
theme_set(theme_bw(base_size=14))
stage.cols = c(seed="springgreen4",juv="blue3",flrg="magenta3")
place.cols = c(Cool ="#175D71" , Warm = "#98463C", Avg = "#D48B47")

sizes = c(seed=1,juv=8,flrg=10)
```

#### 1) survival function parameters
```{r surv_fun}
source("VR.Functions/logit2prob.R")
source("VR.Functions/survival_fun.R")
source("VR.Functions/ParamVals.R")

# look at survival across a range of regional climate values
surv.clim = data.frame(clim = seq(-20,20,by=0.1),
                       seed=NA,
                       juv=NA,
                       flrg=NA)

for(i in 1:nrow(surv.clim)){
  
    surv_params$clim = surv.clim$clim[i]
    s.rates = do.call(survival_fun,surv_params)
    surv.clim[i,2:4]=s.rates
  
}

plot.df = pivot_longer(surv.clim,cols = !clim, names_to = "stage",values_to = "s.rate")

ggplot(plot.df,aes(x=clim,y=s.rate,color=stage))+
  geom_line(linewidth=1.5)+
  scale_color_manual(values=stage.cols) +scale_y_continuous(limits = c(0,1))+
  labs(title="Survival curves")

rm(i)
```

2) growth function parameters
```{r growth_fun}
source("VR.Functions/growth_fun.R")
source("VR.Functions/ParamVals.R")

# look at survival across a range of regional climate values
grow.clim = data.frame(clim = seq(-15,15,by=0.1),
                       seed=NA,
                       juv=NA)

for(i in 1:nrow(grow.clim)){
  
    growth_params$clim = grow.clim$clim[i]
    g.rates = do.call(growth_fun,growth_params)
    grow.clim[i,2:3]=g.rates
  
}

plot.df = pivot_longer(grow.clim,cols = !clim, names_to = "stage",values_to = "g.rate")
ggplot(plot.df,aes(x=clim,y=g.rate,color=stage))+
  geom_line(linewidth=1.5)+
  scale_color_manual(values=stage.cols,labels=c("juv to flrg","seed to juv")) + 
  scale_y_continuous(limits=c(0,1))+
  ggtitle("Growth curves")

rm(i)
```

#### Look at growth fun across range of clim var, asp.eff, asp.mag ####
```{r more_growthFun}
# 
# growth_df = expand.grid(asp.effect = c(-1,0,1),
#                         asp.mag = seq(0,5,by=1),
#                         clim = rnorm(n=1000,sd=3),
#                         seed = NA, juv = NA)
#   
# for(i in 1:nrow(growth_df)){
#   
#   growth_df[i,c('seed','juv')] = growth_fun(sizes = growth_params$sizes,
#                  max.growth.rates = growth_params$max.growth.rates,
#                  min.growth.rates = growth_params$min.growth.rates,
#                  G_2 = growth_params$G_2,
#                  G_3 = growth_params$G_3,
#                  asp.mag = growth_df$asp.mag[i],
#                  asp.effect = growth_df$asp.effect[i],
#                  clim = growth_df$clim[i])
#   
# }
#   
# ggplot(growth_df,aes(x=clim,y=seed,color=as.factor(asp.effect)))+
#   geom_line()+
#   geom_vline(xintercept=0)+
#   facet_wrap(~asp.mag)+
#   theme_bw()+
#   scale_y_continuous(limits = c(0,1))
# 
# ggplot(growth_df,aes(x=clim,y=juv,color=as.factor(asp.effect)))+
#   geom_line()+
#   geom_vline(xintercept=0)+
#   facet_wrap(~asp.mag)+
#   theme_bw()+
#   scale_y_continuous(limits = c(0,1))

```

3) Seed number: To find the optimal seed number, build a mx with the survival and growth rates based on parameters in the two chunks above, then use the optim function to find the optimal seed number that gets lambda ~1.03
```{r seednum}
source("VR.Functions/SG_mx_fun.R")
source("VR.Functions/seednum_fun.R")
source("VR.Functions/ParamVals.R")

growth_params$asp.effect = surv_params$asp.effect = 1
growth_params$asp.mag = surv_params$asp.mag = 0

growth.rates = do.call(growth_fun,growth_params)
surv.rates = do.call(survival_fun,surv_params)

seed_df = expand.grid(move.rate = seq(0,0.5,by=0.25), 
                      aspect.effect = c(-1,0,1),
                      asp.mag = 0:20,
                      seeds=NA)

for(i in 1:nrow(seed_df)){
  
  growth_params$asp.effect = surv_params$asp.effect = seed_df$aspect.effect[i]
  growth_params$asp.mag = surv_params$asp.mag = seed_df$asp.mag[i]

  growth.rates = do.call(growth_fun,growth_params)
  surv.rates = do.call(survival_fun,surv_params)
  
  # Get survival/growth matrix
  sg.mx = sg_mx_fun(surv.rates = surv.rates,growth.rates = growth.rates)
  
  # Run seednum_fun through optim function to fund the seed number that gets a lambda value of around 1.01
  seeds.optim = stats::optim(par = 100, fn = basic_seeds_fun, method = "L-BFGS-B", lower=0, upper=10000000, control = list(factr = 1e-15))
  
  # Save that seed number to the global environment
  seed_df$seeds[i] = round(seeds.optim$par)
  
  sg.mx[1,3] <- round(seeds.optim$par)*P.flr*surv.rates["flrg"]
  Re(eigen(sg.mx)$values[1])

  }

write_csv(seed_df,"VR.functions/fecundity.csv")
```

4) 
```{r subpop.lams}
source("VR.Functions/subpop_mx_fun.R")
source("VR.Functions/ParamVals.R")
fec_df = read_csv("VR.Functions/fecundity.csv")

clim=0
asp.effect = 1
asp.mag = 0
years = 500
pop.vec = rep(100,3)
seeds = mean(fec_df$seeds[fec_df$asp.mag==0])

mx = subpop_mx_fun(surv_params = surv_params, growth_params = growth_params, seeds=seeds);mx
round(mx,4)

Re(eigen(mx)$values[1])
```


```{r subpop.lams}

out_df = data.frame(year = 0:years,
                    s1 = NA,
                    s2 = NA,
                    s3 = NA,
                    total = NA,
                    lam = NA)
out_df[1,2:4] = pop.vec
out_df$total[1] = sum(pop.vec)

for(yr in 1:years){

  pop.vec = mx%*%pop.vec
  row.i = which(out_df$year==yr)

  out_df[row.i,2:4] = pop.vec
  out_df$total[row.i] = sum(pop.vec)
  out_df$lam[row.i] = out_df$total[row.i]/out_df$total[row.i-1]


}

plot_df = out_df %>%
  pivot_longer(!year) %>%
  filter(name!='lam')

ggplot(filter(plot_df,year>-1),aes(x=year,y=value,color=name))+
  geom_point()
```


#### Look at survival fun across range of clim var, asp.eff, asp.mag ####
```{r more_growthFun}
# 
# surv_df = expand.grid(asp.effect = c(-1,0,1),
#                       asp.mag = c(0,1,5),
#                       clim = rnorm(n=1000,sd=3),
#                       seed = NA, juv = NA,flrg=NA)
#   
# for(i in 1:nrow(surv_df)){
#   
#   surv_df[i,c('seed','juv',"flrg")] = 
#     survival_fun(sizes = surv_params$sizes,
#                  max.survival.rates = surv_params$max.survival.rates,
#                  min.survival.rates = surv_params$min.survival.rates,
#                  S_2 = surv_params$S_2,
#                  S_3 = surv_params$S_3,
#                  asp.mag = surv_df$asp.mag[i],
#                  asp.effect = surv_df$asp.effect[i],
#                  clim = surv_df$clim[i])
#   
# }
# 
# surv_df = surv_df %>% 
#   mutate(place.name = case_when(asp.effect==-1~"Cool",
#                                 asp.effect==0~"Avg",
#                                 asp.effect==1~"Warm"))
# 
# ggplot(surv_df,aes(x=clim,y=seed,color=place.name))+
#   geom_line()+
#   geom_vline(xintercept=0)+
#   facet_wrap(~asp.mag)+
#   theme_bw()+
#   scale_color_manual(values=place.cols)
# 
# ggplot(surv_df,aes(x=clim,y=juv,color=as.factor(asp.effect)))+
#   geom_line()+
#   geom_vline(xintercept=0)+
#   facet_wrap(~asp.mag)+
#   theme_bw()
# 
# ggplot(surv_df,aes(x=clim,y=flrg,color=as.factor(asp.effect)))+
#   geom_line()+
#   geom_vline(xintercept=0)+
#   facet_wrap(~asp.mag)+
#   theme_bw()

```

